Dependent Object Types

### 2012-08-30 doc/candidacy/report.tex
summary of context-free syntax of nuObj

%% TODO beg edit
A {\em term} denotes an object or a class. It can be of the following
five forms.
\begin{itemize}
\item A {\em simple name} $x$, which denotes an object.
\item A {\em selection} $t.l$, which can denote either an object or a class.
\item An {\em object creation} \mbox{$\newin{x \larrowc t}{u}$},
which defines a fresh instance $x$ of class $t$.
The scope of this object is the term $u$.
\item 
A {\em class template} $\cls{x: S \sep \tup d}$ where $\tup d$ is a
sequence of {\em definitions} which associate term labels with values
and type labels with types. This acts as a template to construct
objects with the members defined by the definitions.  The name $x$ of
type $S$ stands for ``self'', i.e.\ the object being constructed from
the template.  Its scope is the definition sequence $\tup d$.  A term
or type can refer via $x.l$ \comment{or \mbox{$x.\type \Tsel L$}} to some other
member of that object. No textual sequence constraint applies to such
references; in particular it is possible that a binding refers to
itself or to bindings defined later in the same record.  This
distinguishes our type system from earlier type systems for records
\cite{cardelli-mitchell:records} or modules \cite{harper94type}.
\item 
A {\em mixin composition} $t \withsub{S} u$, which forms a combined
class from the two classes to which $t$ and $u$ evaluate. Here, $S$
is the type of ``self'' in the combined class. 
\end{itemize}
A {\em value} is a simple name or a class template.
A {\em path} $p$ is a name $x$ followed
by a possibly empty sequence of selections, e.g.\ $x . l_1 .; \ldots;
. l_n$.

@@

We let the meta-variable $\preceq$ range over $=$ and $\expand$, and
let $\tbind$ range over $=$, $\expand$, and $<:$.  The name $x$ stands
for ``self''; its type is assumed to be the record type itself. We let
the letter $R$ range over record types.
\item
A {\em compound type} $T \with U$. This type contains all members of
types $T$ and $U$. The subtyping relation for compound types is the
same as the one for intersection types \cite{barendregt-coppo:filter},
but the formation rules are more restrictive.  Where $T$ and $U$ have
a member with the same label, the compound type contains the member
defined in $U$. That member definition must be more specific
(see Section~\ref{sec:type-system}) than the corresponding member
definition in $T$.
\item
A {\em class type} $\cls{x: S \sep \tup D}$, which contains as values
classes that instantiate to objects of type $\obj{x \sep \tup D}$, or
some subtype of it.  $x$ is again the name for ``self''.  It now comes
with an explicit type $S$ which may be different from $\obj{x \sep
\tup D}$. Definitions in $S$ which are missing from $\tup D$ play the
role of abstract members. Such members can be referred to from other
definitions in the class, but they are not defined in the class
itself. Instead, these members must be defined in other classes which
are composed with the class itself in a mixin composition.
Definitions which are present in $\tup D$ but missing in $S$ play in
some sense the role of non-virtual members -- they are not referred to
via ``self'' from inside the class, so overriding them does not change
existing behavior. Definitions present in both $S$ and $\tup D$ play
the role of virtual members.
\end{itemize}
%% TODO end edit

>>>>

The terms in ${\nu}Obj$ denotes objects or classes. They consist of
variables $x$ denoting objects, selections $t.l$, object creations
\mbox{$\newin{x \larrowc t}{u}$} defining an object $x$ of class $t$
whose scope is the term $u$, class templates $\cls{x: S \sep \tup d}$,
and mixing compositions $t \withsub{S} u$ denoting a class of self
type $S$ combined from two classes $t$ and $u$. A value is a variable
or a class template. A path is a variable followed by a possibly empty
sequence of selections.

The types in ${\nu}Obj$ consists of singleton types $p.\type$, type
selections $T\Tsel L$, record types $\obj{x \sep \tup D}$, compound
types $T \with U$, and class types $\cls{x: S \sep \tup D}$.

A class template $\cls{x: S \sep \tup d}$ is closely related to a
translucent sum, except that it needs to be instantiated into an
object while a translucent sum is a module. Like in a translucent sum,
$\tup d$ binds term labels to values and type labels to types. The $x$
of type $S$ stands for the self, i.e.\ the object being constructed
from the template. Its scope is the bindings $\tup d$. Since the self
is explicitly typed, the bindings in class templates can freely refer
to each other (and even to themselves) via $x$, while recall that, in
translucent sums, only later fields can depend on earlier ones.

Contrasting concrete type bindings $\tup d$ used in class template
terms with type declarations $\tup D$ used in record and class types,
the possible type binders in type declarations are type aliases $=$,
new types $\expand$ and abstract types $<:$. Only the first two are
also concrete type binders available in type bindings of class
templates.

\comment{
A {\em term} denotes an object or a class. It can be of the following
five forms.
\begin{itemize}
\item A {\em simple name} $x$, which denotes an object.
\item A {\em selection} $t.l$, which can denote either an object or a class.
\item An {\em object creation} \mbox{$\newin{x \larrowc t}{u}$},
which defines a fresh instance $x$ of class $t$.
The scope of this object is the term $u$.
\item 
A {\em class template} $\cls{x: S \sep \tup d}$ where $\tup d$ is a
sequence of {\em definitions} which associate term labels with values
and type labels with types. This acts as a template to construct
objects with the members defined by the definitions.  The name $x$ of
type $S$ stands for ``self'', i.e.\ the object being constructed from
the template.  Its scope is the definition sequence $\tup d$.  A term
or type can refer via $x.l$ \comment{or \mbox{$x.\type \Tsel L$}} to some other
member of that object. No textual sequence constraint applies to such
references; in particular it is possible that a binding refers to
itself or to bindings defined later in the same record.  This
distinguishes our type system from earlier type systems for records
\cite{cardelli-mitchell:records} or modules \cite{harper94type}.
\item 
A {\em mixin composition} $t \withsub{S} u$, which forms a combined
class from the two classes to which $t$ and $u$ evaluate. Here, $S$
is the type of ``self'' in the combined class. 
\end{itemize}
A {\em value} is a simple name or a class template.
A {\em path} $p$ is a name $x$ followed
by a possibly empty sequence of selections, e.g.\ $x . l_1 .; \ldots;
. l_n$.

@@

We let the meta-variable $\preceq$ range over $=$ and $\expand$, and
let $\tbind$ range over $=$, $\expand$, and $<:$.  The name $x$ stands
for ``self''; its type is assumed to be the record type itself. We let
the letter $R$ range over record types.
\item
A {\em compound type} $T \with U$. This type contains all members of
types $T$ and $U$. The subtyping relation for compound types is the
same as the one for intersection types \cite{barendregt-coppo:filter},
but the formation rules are more restrictive.  Where $T$ and $U$ have
a member with the same label, the compound type contains the member
defined in $U$. That member definition must be more specific
(see Section~\ref{sec:type-system}) than the corresponding member
definition in $T$.
\item
A {\em class type} $\cls{x: S \sep \tup D}$, which contains as values
classes that instantiate to objects of type $\obj{x \sep \tup D}$, or
some subtype of it.  $x$ is again the name for ``self''.  It now comes
with an explicit type $S$ which may be different from $\obj{x \sep
\tup D}$. Definitions in $S$ which are missing from $\tup D$ play the
role of abstract members. Such members can be referred to from other
definitions in the class, but they are not defined in the class
itself. Instead, these members must be defined in other classes which
are composed with the class itself in a mixin composition.
Definitions which are present in $\tup D$ but missing in $S$ play in
some sense the role of non-virtual members -- they are not referred to
via ``self'' from inside the class, so overriding them does not change
existing behavior. Definitions present in both $S$ and $\tup D$ play
the role of virtual members.
\end{itemize}
}
